---
title: "Project 2D: Appendices"
author: "Paul Motten"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: cosmo
    highlight: textmate
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float: false
    
---

## Author's Notes:

These are bits I wrote in the process of completing the assignment which did not make the final cut.  Some of them don't work, some don't show what they're supposed to show, and others are simply intermediate steps in developing the visualizations used in earlier pages.  Some of the code, however, was recycled into the final versions.  

The most recent versions of exploratory visualizations are included here for future reference.


## Packages Used

```{r loadpackages, collapse = TRUE}
library(tidyverse)
library(maps)
library(knitr)
if (!require("gganimate")) install.packages("gganimate")
library(gganimate)
if (!require("transformr")) install.packages("transformr")
library(transformr)
if (!require("ggbeeswarm")) install.packages("ggbeeswarm")
library(ggbeeswarm)
if (!require("plotly")) install.packages("plotly")
library(plotly)
if (!require("kableExtra")) install.packages("kableExtra")
library(kableExtra)
if (!require("DT")) install.packages("DT")
library(DT)
if(!require("waffle")){install.packages("waffle")}
library(waffle)
```

## Appendix A

###   Notes

 ... For `rename()`: `<tidy-select>` Use `new_name = old_name`

 ... `ifelse()` args:  `(test, yes, no)`


## Appendix B

```{r}

#
# ###  Compare the QUALITY of Disney+ to other offerings

strmtv_num <- strmtv_long %>%
  mutate(
    Rotten = as.numeric(sub(x = strmtv_long$`Rotten Tomatoes`, "/.*", "")),
    IMDb_num = ifelse(is.na(IMDb), NA, as.numeric(sub(x = strmtv_long$IMDb, "/.*", "")) * 10),
    Critics = ifelse(is.na(IMDb), Rotten, (Rotten + IMDb_num) / 2)
  )

ggplot(strmtv_num, aes(Service, Critics, fill = Service)) +
  geom_violin(bw = 2, scale = "count") +
  scale_fill_manual(
      breaks = c("Disney+", "Hulu", "Netflix", "Prime Video"),
      values = c("darkorchid2", "green3", "firebrick2", "dodgerblue")) +
  scale_y_continuous(
    n.breaks = 5, 
    minor_breaks = seq(0, 100, 5),
    limits = c(0, 100)) +
  labs(title = "Quality of TV Shows on Selected Streaming Services",
       x = "Number of Programs (width of plot)",
       y = "Average % Positive Critical Reviews")

```

## Appendix C

```{r loaddata, collapse = TRUE}
bills <- read_csv("../dataraw/billionaires_2021_10_31.csv")

glimpse(bills)
is_tibble(bills)

ipeds <- read_csv("https://assets.datacamp.com/production/repositories/1942/datasets/18a000cf70d2fe999c6a6f2b28a7dc9813730e74/ipeds.csv")
glimpse(ipeds)

teen <- readRDS("../dataprocessed/CDCteenbirthrate.rds")
str(teen)

teenNC <- readRDS("../dataprocessed/CDCteenbirthrateNC.rds") %>% 
  mutate(subregion = tolower(county), # to match county map data
         year = as.integer(year)) # to make the animation labels whole numbers
str(teenNC)

### County Data from Maps Package
countymapNC <- map_data("county") %>% 
  filter(region == "north carolina") %>% 
  select(-region) # not needed since it will be all NC data now
str(countymapNC)

```

```{r exploration, collapse = TRUE}

# bills[bills == "$121B"] <- NA  # Check that is.na check is checking correctly.  
#   ###
#   ###  Re-run read_cvs to refresh dataframe after testing
#   ###
#
for (i in 1:ncol(bills)) {
  print(which(is.na(bills[ ,i])))
}


bills_ex <- bills %>%
  group_by(Industry) %>%
  summarise(Sum = sum(Total_Net_Worth_Bil))

bills_red <- bills %>%
  mutate(
    Industry = ifelse(
        bills$Industry == "Technology" |
        bills$Industry == "Industrial" |
        bills$Industry == "Finance" |
        bills$Industry == "Diversified"|
        bills$Industry == "Consumer",
      Industry,
      "Other"),
    LastChange_Prop = (LastChange_Bil + Total_Net_Worth_Bil) / Total_Net_Worth_Bil,
    LastChange_Coef = (LastChange_Bil + mean(Total_Net_Worth_Bil)) / mean(Total_Net_Worth_Bil)
  ) %>%
  rename(`Net Worth (Billions)` = Total_Net_Worth_Bil,
         `Change YTD` =  YTDChange_Bil)
glimpse(bills_red)

bills_rex <- bills_red %>%
  group_by(Industry) %>%
  summarise(Sum = sum(`Net Worth (Billions)`))


ggplot(bills, aes(LastChange_Bil, Total_Net_Worth_Bil, color = Industry)) +
    geom_point() #+
#    scale_x_continuous(trans = "log2") +
#    scale_y_continuous(trans = "log2")
```

```{r exploratory plotting}


ggplot(bills_red, aes(x = Industry, fill = Industry)) +
    geom_bar()

ggplot(bills, aes(x = Industry, fill = Industry)) +
    geom_bar() +
    coord_flip()

### Might not be graphing what I think they're graphing?

ggplot(bills_ex, aes(Industry, Sum, fill = Industry)) +
    stat_summary(geom="bar", position = "stack") +
    xlab("Sector") +
    coord_flip()

ggplot(bills_rex, aes(Industry, Sum, fill = Industry)) +
    stat_summary(geom="bar", position = "stack") +
    xlab("Sector")

#
##  ***** Non-functional code, do not use ******
##  plot(bills, aes(LastChange_Bil, Total_Net_Worth_Bil, color = Country)) +
#     geom_point() +
#     scale_x_continuous(n.breaks = 10) +
#     scale_y_continuous(n.breaks = 10)
#


```
```{r}

jitter <- position_jitter(width = 0.2, height = 0, seed = 5)

ggplot(bills_red,
       aes(
         x = 1,
         y = LastChange_Prop,
         color = Industry,
         size =  `Net Worth (Billions)`
       )) +
  geom_point(position = jitter, alpha = 0.8) +
  scale_size(range = c(0, 20)) +
  theme(title = element_text("YTD Change in wealth relative to current fortune"))

ggplot(bills_red,
       aes(
         x = 1,
         y = LastChange_Coef,
         color = Industry,
         size =  `Net Worth (Billions)`
       )) +
  geom_point(position = jitter) +
  scale_size(range = c(0, 20)) +
  theme(title = element_text("Absolute change in wealth YTD"))


```


```{r teendelta}

teenNC_wide <- teenNC %>%
  select(-c("LCL", "UCL")) %>%
  pivot_wider(names_from = year, names_prefix = "births_", values_from = birth_rate) %>%
  mutate(births_delta = births_2018 - births_2003,
         birth_incr = ifelse(births_delta >= 0, births_delta, NA))
  
```

```{r mapjoin} 
mapdata <- countymapNC %>% 
  full_join(teenNC_wide, by = "subregion")
```

```{r}
ipedsNC <- ipeds %>%
  filter (
    lat < 36.5,
    lng < -75.554359,
    lat > 33.888967,
    lng > -84.321449,
    sector_label == "Public"
  )

print(ipedsNC)
```

```{r}
mapdata %>%
#  filter(year == "2003") %>%                       fill = birth_rate),
   ggplot() +
   geom_polygon(aes(x = long, 
                    y = lat, 
                    group = group, 
                    fill = births_delta),
                color = "black") +
  coord_map() +
  scale_fill_gradient2(low = "green", mid = "white", high = "red", midpoint = 0) +
  geom_point(data = ipedsNC, 
             aes(lng, lat, color = "Location of UNC Campuses"),
             size = 1.5) +
    theme(plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) +
    scale_colour_manual(values = "black", na.value = "black") +
    labs(title = "Change in Teen Birth Rates in North Carolina",
         subtitle = "From 2003 to 2018, per 1,000 females in age group 15–19 years",     #
         fill = "Teen Birth Rate",
         color = NULL,
         caption = "Data Source: Centers for Disease Control and Prevention") 

```


```{r}
teen_vis <- teen %>%
  filter(state == "North Carolina",
         county == "Watauga")

ggplot(teen_vis, aes(year, birth_rate)) +
    stat_smooth(alpha = 0.2) +
  geom_line() +
  scale_y_continuous(limits = c(0, 12)) +
  labs(title = "Change in Teen Birth Rates, Watauga County NC",
       subtitle = "per 1,000 females in age group 15–19 years",
       caption = "Data Source: Centers for Disease Control and Prevention") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

teen_vis2 <- teen %>%
  filter(state == "North Carolina") %>%
  group_by(county)

ggplot(teen_vis2, aes(year, birth_rate, color = county)) +
  geom_line() +
  labs(title = "Change in Teen Birth Rates NC Counties",
       subtitle = "per 1,000 females in age group 15–19 years",
       caption = "Data Source: Centers for Disease Control and Prevention") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")

```


```{r}
sessionInfo()
```